---
title: "Mod4_DataVizR"
author: "Alan Thomas Joshy"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install and load the tidyverse package
```{r}
#install.packages("tidyverse")
library(tidyverse)
```
#Load Data in
```{r}
data(mpg)
```

# Creating the first plot using ggplot
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy))
```


playing around with different "mappings" of ggplot such as colour, size and gradient

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy, colour = class, alpha = fl))
```

#R gives us a warning against using alpha for a discrete variable as I have in the above trial plot.

# Using facets to highlight subsets of Data
```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)
```

#facet_wrap() is only used for discrete variables

facet_grid() can be used when a combination of two variables have to be plotted.

```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

#Excercise: "nrow" and "ncol" refers to the number of rows and columns in the plot

#Fitting Lines on Plots
```{r}
ggplot(mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
geom_smooth can be used to add smooth lines onto plots, instead 

#we can also have the points on the plot along with the smooth line
```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

```

#we can also have multiple lines on the plot to denote different "stuff"
```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, colour = drv, linetype = drv))
```

#we can use the group aesthetic to group categorical variables
```{r}
ggplot(mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv, color = drv),
             show.legend = TRUE,
             )
```

