---
title: "Mod4_Session 1"
author: "Alan Thomas Joshy"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install and load the tidyverse package
```{r}
#install.packages("tidyverse")
library(tidyverse)
```
#Load Data in
```{r}
data(mpg)
```

# Creating the first plot using ggplot
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy))
```


playing around with different "mappings" of ggplot such as colour, size and gradient

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y= hwy, colour = class, alpha = fl))
```

#R gives us a warning against using alpha for a discrete variable as I have in the above trial plot.

# Using facets to highlight subsets of Data
```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~class, nrow = 2)
```

#facet_wrap() is only used for discrete variables

facet_grid() can be used when a combination of two variables have to be plotted.

```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ cyl)
```

#Excercise: "nrow" and "ncol" refers to the number of rows and columns in the plot

#Fitting Lines on Plots
```{r}
ggplot(mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```
geom_smooth can be used to add smooth lines onto plots, instead 

#we can also have the points on the plot along with the smooth line
```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = drv)) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))

```


#we can use the group aesthetic to group categorical variables
```{r}
ggplot(mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv, color = drv),
             show.legend = TRUE,
             )
```

#we can also have multiple lines on the plot to denote different "stuff"
```{r}
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, colour = drv, linetype = drv))
```

# there is a better way to code the above plot in a more efficient way and one that prevents uneccesary mistakes:
```{r}
ggplot(mpg, mapping = aes(x = displ, y = hwy, colour = drv)) + 
  geom_point(mapping = aes()) + 
  geom_smooth()
``` 
using mapping under ggplot in the geom_ point and geom_smooth will allow us to make changes to the different layers.
```{r}
ggplot(mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(colour = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = TRUE)
```

#Transformations and Stats:

```{r}
data("diamonds")
```

#Bar graph of Diamonds data
```{r}
ggplot(diamonds,) +
  geom_bar(mapping = aes(x = cut))
```

We are going to change the count of diamonds which is an r calculated stat to identity, which is a raw value of the variable.
```{r}
demo <-  tribble(
  ~cut, ~freq,
  "Fair", 1610,
  "Good", 4906,
  "Very Good", 12082,
  "Premium", 13791,
  "Ideal", 21551
)
demo
```
```{r}
ggplot(data = demo) + 
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, y = after_stat(prop), group = 1))
```

#we can plot the statistical details for transparency
```{r}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max, 
    fun = median
  )
```

#aesthetics like colour and fill on geom_bar:
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, colour = cut))
```
colour gives outline vs fill gives fully coloures bars


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

#Position adjustments:
Identity
```{r}
#position = identity
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(aes(colour = clarity), position = "identity")

#position = fill 
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(aes(colour = clarity), position = "fill")

#position = dodge
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) +
  geom_bar(aes(colour = clarity), position = "dodge")
```

#jitter position:
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "jitter")
```

