---
title: "Module4_Workshop 3"
author: "Alan Thomas Joshy"
date: "2024-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Initially load the package tidyverse
```{r}
library(tidyverse)
```

Trials at Piping (%>%) (|>):

```{r}
data <-  table1
```


actual piping:
```{r}
table1 |>
  mutate(rate = cases / population * 10000)
```

this piping achieves the same result as the following code using '$' signs to denote variables from the dataset- table1 

```{r}
mutate(table1, rate = table1$cases / table1$population * 10000 )
```
Pivoting in order to tidy up data:
pivot_longer used for decreasing number of columns and increase number of rows
pivot_wider used for the opposite:
```{r}
billboard |>
pivot_longer(
  cols = !c(artist, track, date.entered),
  names_to = "Week",
  values_to = "Rank",
  values_drop_na = TRUE
  )
```
values_drop_na drops or ignores unknown values.


```{r}
df <- tribble(
  ~id, ~bp1, ~bp2, ~bp3,
  "A", 100, 120, 140,
  "B", 140, 115, 150,
  "C", 120, 125, 175
)
```


```{r}
df |> 
  pivot_longer(
    cols = c(bp1,bp2,bp3),
    names_to = "measurement",
    values_to = "value"
  )
```

