---
title: "Workshop4_GISinR"
author: "Alan Thomas Joshy"
date: "2024-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load tidyverse into the library for the session
```{r}
library("tidyverse")
```

Install and load the new packages required
```{r}
#install.packages("sf")
#install.packages("terra")
#install.packages("tmap")

#load into library
library(sf) #simple features
library(terra) #for raster
library(tmap) #Thematic maps are geographical maps in which spatial data distributions are visualized

```
```{r}
getwd()
```


We can load our copepod data using a package called readr, which is part of the tidyverse. readr > base R functions for data consistency
```{r}
#working directory default set to folder containing .rmd file, so when loading data specify to go back up a folder using "../"
library(readr)
dat <- read_csv("../data/data-for-course/copepods_raw.csv")
dat
```


First we visualize the data to get an initial idea on it, and we'll use ggplot initially and tmap
```{r}
library(ggplot2)
ggplot(dat) +
  aes(x = longitude, y = latitude, colour = richness_raw) +
  geom_point()
```

This is simply a nice and easy way to look at the data but it does not have critical characteristics needed for a map 

Now we will visualize the richness data in a non-spatial way
```{r}
ggplot(dat, aes(x = latitude, y = richness_raw)) +
  stat_smooth() +
  geom_point()
```

Now plot some different variables from the data

```{r}
dat
```
```{r}
ggplot2 :: ggplot(dat, aes(x = region)) + #not really a good plot, doent give any info on data...
  geom_bar()
```


```{r}
#calculate mean richness for region
dat_sum <- dat |>
  group_by(region) |>   #group by region
  summarise(mean_richness = mean(richness_raw), #get mean richness
  sd_richness = sd(richness_raw))  #calculate standard deviation 

#plot the mean richness against region graph
ggplot(dat_sum, aes(x = region, y = mean_richness)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean_richness - sd_richness, ymax = mean_richness + sd_richness),width = 0.2) + 
  labs(title = "Mean Richness by Region", x = "Regions", y = "Mean Species Richness")

  
```

